\textbf{International Transport costs:\\New Findings from modeling additive costs}\\Online Appendix (Not for Publication)

### 2021-02-27 redaction/revision_JEGeo/revised_article/DHP_JoEG_rev1.tex
changement rédaction papier

From Equation (\ref{eq:tt1}), we can identify a first polar case $\beta=0$. Shipping charges are proportional to the fas price, such that $p=\widetilde{p}(1+X)$: Only ad-valorem costs apply. At the other extreme, the case $\beta=1$ corresponds to shipping costs being independent of the unit price, with the cif price the sum of the unit price and the shipping cost: Transport costs are fully additive.
This suggests that, the closer $\beta$ to 1, the more prevalent additive costs in total transport costs. One can analytically show that $\beta$ is indeed equal to the share of additive costs in total transport costs.

@@

\begin{eqnarray}
&& p = \tau \widetilde{p}+ t \label{eq:tt0} \\
\Leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
\end{eqnarray}

\noindent with $\tau \geq 1$ the ad-valorem cost ($\tau =1$ meaning no ad-valorem cost) and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})). Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}. Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.}
\begin{eqnarray}
\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau - 1  +\frac{t}{\tilde{p}}} \\
&=& \frac{-t}{\tau -1+\widetilde{t}} \label{eq:tt2}
\end{eqnarray}

@@

\paragraph{Obtaining the estimated equation} Let us denote $i$ the origin country, and $k$ the product (at the 5-digit level in our benchmark estimate), for each flow that is year- and transport-mode specific. For sake of reading convenience though, we skip the year and transport-mode dimensions in the notations. Transforming the above equation (\ref{eq:base}) as the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation:

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} -1 +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
\end{equation}

We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i\times \tau_k -1 \simeq (\tau_i -1)+ (\tau_k-1)$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
We also provide a robustness analysis to the separability assumption. Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries. See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}

\begin{eqnarray}
\tau_{ik} &=&  \tau_{i} \tau_{k} \label{eq:ad-valorem}\\
t_{ik} &=& t_{i} + t_{k} \label{eq:add}
\end{eqnarray}

@@

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\tau_{i}\tau_{k} -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
\end{equation*}

The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the cif price $p$ cannot be lower than the fas price ($p_{ik}>\widetilde{p}_{ik}$).
Taking into account this constraint in the estimate implies that the error term should be always positive.
We ensure that this constraint is fulfilled by specifying the error term as follows:

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left(\tau_{i}\tau_{k} -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
\end{equation*}
\noindent where $\epsilon_{ik}$ follows a normal law centered on 0.
Considered in logs, the above equation becomes:

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i}\tau_{k}+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
\end{equation}

Equation (\ref{eq:equation0}) is non-linear. It cannot be estimated using standard linear estimators.
All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations.
The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at a more disaggregated product level ($k=5$ in our benchmark estimates). This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the
same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
This would not solve the problem, however, for three main complementary reasons.
First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}

This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\tau_{i} \tau_{s(k)}+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:model_IetA}
\end{equation}
where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, strictly speaking, the exact equation that we estimate is the following:
$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode).
For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5\% of the upper and lower tails of the distribution in the regression variables.
%JH: TO KEEP IN MIND: one day, we could be asked to winsorize tails, instead of cutting them...
These cut-offs are aimed at eliminating reporting or coding errors.
We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
  \medskip

>>>>

From Equation (\ref{eq:tt1}), we can identify a first polar case where $\beta=0$. Shipping charges are proportional to the fas price, such that $p=\widetilde{p}(1+X)$: Only ad-valorem costs apply. At the other extreme, the case $\beta=1$ corresponds to shipping costs being independent of the unit price, with the cif price the sum of the unit price and the shipping cost: Transport costs are fully additive.
This suggests that, the closer $\beta$ to 1, the more prevalent additive costs in total transport costs. One can analytically show that $\beta$ is indeed equal to the share of additive costs in total transport costs.

@@

\begin{eqnarray}
&& p = (1+\tau) \widetilde{p}+ t \label{eq:tt0} \\
\Leftrightarrow && p = (1+\tau) \widetilde{p}+ \widetilde{t}\widetilde{p}  \label{eq:base}
\end{eqnarray}

\noindent with $\tau \geq 0$ the ad-valorem cost and $t>0 $ the additive shipping charge per kg transported (Equation (\ref{eq:tt0})). Notice that we can rewrite the additive cost as a function of the cost per USD shipped $\widetilde{t} \equiv \frac{t}{\widetilde{p}}$, as in Equation \ref{eq:base}. Starting from Equation (\ref{eq:base}), the elasticity of transport costs to the unit price $\widetilde{p}$ can be shown to be equal to:\footnote{See Appendix XXX for a formal demonstration.}
\begin{eqnarray}
\frac{\partial TC}{\partial \widetilde{p}} \frac{\tilde{p}}{TC}&=& \frac{-\frac{t}{\tilde{p}}}{\tau   +\frac{t}{\tilde{p}}} \notag \\
&=& \frac{-t}{\tau +\widetilde{t}} \label{eq:tt2}
\end{eqnarray}

@@

\paragraph{Obtaining the estimated equation} Let us denote $i$ the origin country, and $k$ the product (at the 5-digit level in our benchmark estimate), for each year-mode specific flow. For sake of reading convenience, we skip the year and transport-mode dimensions in the notations. Transforming the above equation (\ref{eq:base}) as the percentage change in prices induced by transportation (and get rid of the kg unit), we get the following baseline specification underlying our estimation:

\begin{equation}
\frac{p_{ik}}{\widetilde{p}_{ik}} -1 = \tau_{ik} +\frac{t_{ik}}{ \widetilde{p}_{ik}} \label{eq:base_estimee}
\end{equation}

We follow \cite{Irrazabal_2015} by considering that \textit{i)} both ad-valorem and additive costs are separable between the origin country ($i$) and the product ($k$) dimensions, and \textit{ii)} this separability is in a multiplicative way for the former and an additive way for the latter.
In other words, $\tau_{ik}$ and $t_{ik}$ from Equation (\ref{eq:base_estimee}) are written as:\footnote{Notice that, given the magnitude or order of transport costs, assuming an additive or a multiplicative form for country/product fixed effects does not make a substantial difference since, for small values (as we generally obtain), we have $\tau_i+ \tau_k \simeq (1+\tau_i)\times(1+\tau_k) -1$ and $t_i+t_k\simeq (1+t_i)\times(1+t_k)-1$.
We also provide a robustness analysis to the separability assumption. Given the much higher number of fixed effects to estimate, we run this robustness check on a sub-sample of goods/origin countries. See Section \ref{sec:robustness} for a detailed presentation of this robustness check.}

\begin{eqnarray}
1+\tau_{ik} &=& (1+ \tau_{i})(1+ \tau_{k}) \label{eq:ad-valorem}\\
t_{ik} &=& t_{i} + t_{k} \label{eq:add}
\end{eqnarray}

@@

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =(1+\tau_{i})(1+\tau_{k}) -1 +\frac{t_{i} + t_{k}}{ \widetilde{p}_{ik}} \label{eq:theory_equation}
\end{equation*}

The ratio $\frac{p_{ik}}{\widetilde{p}_{ik}}$ has a lower bound of one, since by construction, the import price $p$ cannot be lower than the export price ($p_{ik}\geq\widetilde{p}_{ik}$). Taking into account this constraint in the estimate implies that the error term should be always positive.
We ensure that this constraint is fulfilled by specifying the error term as follows:

\begin{equation*}
\frac{p_{ik}}{\widetilde{p}_{ik}}-1 =\left((1+\tau_{i})(1+\tau_{k}) -1+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}} \right)\times \exp(\epsilon_{ik})
\end{equation*}
\noindent where $\epsilon_{ik}$ follows a normal law centered on 0.
Considered in logs, the above equation becomes:

\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left((1+\tau_{i})(1+\tau_{k})+\frac{t_{i} + t_{k}}{\widetilde{p}_{ik}}-1 \right) + \epsilon_{ik} \label{eq:equation0}
\end{equation}

Since Equation (\ref{eq:equation0}) is non-linear, it cannot be estimated using standard linear estimators. All estimates are thus performed using non-linear least squares.\footnote{The basis of the method is to approximate the model by a linear one and to refine the parameters by successive iterations. The intuitive criterion for convergence is that the sum of squares of residuals does not increase from one iteration to the next.
See \cite{Woolridge-Book-2001} for more details.} Yet, the use of a non-linear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time.
Confronted with this trade-off, we estimate international transport costs at the 3-digit level as our benchmark classification, even though data series are available at a more disaggregated product level ($k=5$ in our benchmark estimates). This amounts to making the additional assumption, that all 5-digit products $k$ in a 3-digit sector $s$ share the same structure of costs.\footnote{One way to gauge the relevance of this assumption is to provide a decomposition variance exercise on the observed cif-fas price.
As reported in Appendix \ref{app:decomp_variance}, the share of the observed variance that is accounted for by the between-sectors ($s$) variance is roughly similar to the between-product ($k$) variance.}$^{,}$\footnote{One may object that we could preserve the estimation at the 5-digit level by running an OLS estimation on the equation taken in level, i.e.
on the basis of Equation (\ref{eq:base_estimee}), specifying the error term additively.
This would not solve the problem though, for three main complementary reasons.
First, at the 5-digit level the number of fixed effects to include in the estimation would be more than 430,000 (with 216 countries and 2,029 products), making the estimation computationally extremely burdensome, even in OLS.
Imposing Equations (\ref{eq:ad-valorem}) and (\ref{eq:add}) to reduce the number of fixed effects would drive us back to the non-linearity issue for fixed effects.
Second, making the error term (specified as following a normal law centered on 0) enter Equation (\ref{eq:base_estimee}) additively implies negative values for some of the estimated residuals $\widehat{\epsilon}_{ik}$, which is inconsistent with the constraint that $\frac{p_{ik}}{\widetilde{p}_{ik}}>1$.
Third, estimating the equation in level does not eliminate the fact that it is non-linear by nature, as long as there are additive costs to estimate (see Equation (\ref{eq:base_estimee}) for $t_{ik} \neq 0$).}

This drives us to estimate a modified version of Equation (\ref{eq:equation0}), specified as:
\begin{equation}
\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left((1+\tau_{i})(1+ \tau_{s(k)})-1+\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}} \right) + \epsilon_{ik} \label{eq:model_IetA}
\end{equation}
where $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$ and $t_{s(k)}$ are the parameters to be estimated, i.e., fixed effects specific to each origin country $i$ and sector $s$ (at the 3-digit classification level), and $\epsilon_{ik}$ the residual centered on 0.\footnote{Note that, strictly speaking, the exact equation that we estimate is the following:
$$\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\sum_i \alpha_i^\tau \mathbbm{1}_i \times \sum_{s(k)}\alpha_{s(k)}^\tau \mathbbm{1}_{s(k)}+ \frac{\sum_i\alpha_i^t\mathbbm{1}_i + \sum_{s(k)}\alpha^t_{s(k)}\mathbbm{1}_{s(k)}}{\widetilde{p}_{ik}}-1\right)+\epsilon_{ik} $$
After proceeding to the estimation, we uncover the estimated values of the transport costs components $\tau_{i}$, $\tau_{s(k)}$, $t_{i}$, $t_{s(k)}$ from the obtained coefficients $\alpha^{x}_{z}$ for $x=\tau,t$ and $z=i,s(k)$ (by year and transport mode).
For reading convenience, we adopt a lighter expression of this equation by resorting to the notations of $\tau_i,\tau_{s(k)}$ and $t_i$, $t_{s(k)}$ as a short-cut.}  To eliminate the potential influence of outliers, we exclude 5\% of the upper and lower tails of the distribution in the regression variables.
%JH: TO KEEP IN MIND: one day, we could be asked to winsorize tails, instead of cutting them...
These cut-offs are aimed at eliminating reporting or coding errors.
We estimate Equation (\ref{eq:model_IetA}) for each year over the period 1974-2013, for each of the transportation modes reported (air or vessel), on a sectoral-origin country basis ($i,s$).
Depending on the year considered, this leaves us with around 800 fixed effects to estimate by transport mode at the 3-digit level.
  \medskip
